---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Emily R. H. Kelly"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
```{r}
setwd("~/R/R - Datascripts/Assignment 1 and 2/Ihavenoideawhatimdoing-master (1)/Ihavenoideawhatimdoing-master")

library(ggplot2);library(pastecs);library(rmarkdown);library(tidyr);library(plyr);
library(dplyr);library(QuantPsyc);library(VIF);library(stringr);library(tidytext);library(DHARMa)
library(lme4);library(lmerTest);library(psych);library(MuMIn);library(tidyverse);library(magrittr);library(Metrics);library(caret);library(simr)
#Load dataset from part 1
ASDdata = read.delim('ASD_data.csv', sep = ',')

----------------------------------------------------------------
#Model without interaction
#Fit newlm3 model from assignment P3. 
Bmodel3 <- lmer(CHI_MLU ~ Diagnosis + Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=ASDdata, REML = FALSE)
summary(Bmodel3)


#Asses power of main effect. nsim = amount of simulations.
#Power for predictor = 100.0% (92.89, 100.0)
#Effect size for Visit is 0,091
powerV <- powerSim(Bmodel3,fixed("Visit"), nsim=50)
powerV

#Power for diagnosis. 
#Power for predictor = 36% (29.99, 58.75)
#Effect size for diagnosis is -0,32.
powerV_Diag <- powerSim(Bmodel3,fixed("Diagnosis"),nsim=50)
powerV_Diag
summary(Bmodel3)

#Effect size for MOT_MLU is 0.39
#95% confidence interval 100% (92.89, 100)
powerV_MOTMLU <- powerSim(Bmodel3, fixed("MOT_MLU"), nsim = 50)
powerV_MOTMLU

#Effect size for ADOS1 is -0,025
#Power for predictor = 34.00% (21.21, 48.77) 
powerV_ADOS1 <- powerSim(Bmodel3, fixed("ADOS1"), nsim = 50)
powerV_ADOS1

#Effect size for tokens_MOT is -0,00029
#Power for predictor = 100% (92.89, 100)
powerV_tokens_MOT <- powerSim(Bmodel3, fixed("tokens_MOT"), nsim = 50)
powerV_tokens_MOT

#Effect size for tokens_CHI is 0,0017
#Power for predictor = 100% (92.89, 100)
powerV_tokens_CHI <- powerSim(Bmodel3, fixed("tokens_CHI"), nsim = 50)
powerV_tokens_CHI

powerCurveV = powerCurve(Bmodel3, fixed("Visit"),along="Child.ID", nsim=50)
powerCurveV
plot(powerCurveV)



----------------------------------------------------------------
#Model with interactions.
#Fit newml3 model from assignment P3. 
Bmodel3_int <- lmer(CHI_MLU ~ Diagnosis * Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=ASDdata, REML = FALSE)
summary(Bmodel3_int)

#Power for predictor = 100% 
#effectsize = 0.162
powerVInt_Interact <- powerSim(Bmodel3_int, fixed("Diagnosis:Visit"),nsim=50)
powerVInt_Interact


#Effect size for MOT_MLU is 0,38
#Power for predictor = 100% (92.89, 100.0)
powerVInt_MOTMLU <- powerSim(Bmodel3_int, fixed("MOT_MLU"), nsim = 50)
powerVInt_MOTMLU

#Effect size for ADOS1 is -0,029
#Power for predictor = 36,00% (22.92, 50.81)
powerVInt_ADOS1 <- powerSim(Bmodel3_int, fixed("ADOS1"), nsim = 50)
powerVInt_ADOS1

#Effect size for tokens_MOT is  -0,00027
#Power for predictor =  100% (69,15, 100)
powerVInt_tokens_MOT <- powerSim(Bmodel3_int, fixed("tokens_MOT"), nsim = 50)
powerVInt_tokens_MOT

#Effect size for tokens_CHI is 0,0017
#Power for predictor = 100% (92.89, 100)
powerVInt_tokens_CHI <- powerSim(Bmodel3_int, fixed("tokens_CHI"), nsim = 50)
powerVInt_tokens_CHI



```

- Report the power analysis and comment on what you can (or cannot) use its estimates for.

I can use my estimates for performing my poweranalysis. 

 
### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
```{r}
#To perform a more conservative power analysis i would probably pick lower effect sizes.
#I would pick the minium effect sizes that i found in exercise 1 from running a simulation on each fixed effect. I did this for my model with and without interaction.

#More conservatuve effectsize, minus with 1 or 2 SD.

#Without interaction______________________________________________________________________________
summary(Bmodel3, digits = 3)
fixef(Bmodel3)["Visit"] <- 0.091-0.185
fixef(Bmodel3)["DiagnosisTD"] <- -0.32-0.202
fixef(Bmodel3)["MOT_MLU"] <- 0.39-0.046
fixef(Bmodel3)["ADOS1"] <- -0.025-0.014
fixef(Bmodel3)["tokens_MOT"] <- -0.00029-0.00005943
fixef(Bmodel3)["tokens_CHI"] <- 0.0017-0.0001157

Bmodel3_int <- lmer(CHI_MLU ~ Diagnosis + Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=ASDdata, REML = FALSE)
summary(Bmodel3_int)

#Visit
#we see that for reaching a power of 80 for our effect we need 20< participants
powerCurveV = powerCurve(Bmodel3, fixed("Visit"),along="Child.ID", nsim=10)
powerCurveV
plot(powerCurveV)

#Diagnosis
#We see that we need more than 60 participants
powerCurveV_diag= powerCurve(Bmodel3, fixed("Diagnosis"),along="Child.ID", nsim=50)
powerCurveV_diag
plot(powerCurveV_diag)

#MOT MLU
powerCurveV_MOT= powerCurve(Bmodel3, fixed("MOT_MLU"),along="Child.ID", nsim=50)
powerCurveV_MOT
plot(powerCurveV_MOT)

#ADOS1
powerCurveV_ADOS1= powerCurve(Bmodel3, fixed("ADOS1"),along="Child.ID", nsim=50)
powerCurveV_ADOS1
plot(powerCurveV_ADOS1)

#tokens_MOT
powerCurveV_TMOT= powerCurve(Bmodel3, fixed("tokens_MOT"),along="Child.ID", nsim=50)
powerCurveV_TMOT
plot(powerCurveV_TMOT)

#tokens_CHI
powerCurveV_TCHI= powerCurve(Bmodel3, fixed("tokens_CHI"),along="Child.ID", nsim=50)
powerCurveV_TCHI
plot(powerCurveV_TCHI)



#With interaction, replace fixed() in curve with fixed effect for interaction. __________________ 
summary(Bmodel3_int)
fixef(Bmodel3_int)["MOT_MLU"] <- 0.38-0.045
fixef(Bmodel3_int)["ADOS1"] <- -0.029-0.014
fixef(Bmodel3_int)["tokens_MOT"] <- -0.00027-0.00005814
fixef(Bmodel3_int)["tokens_CHI"] <- 0.0017-0.0001134
fixef(Bmodel3_int)["DiagnosisTD:Visit"] <- 0.162-0.026

Bmodel3_int <- lmer(CHI_MLU ~ Diagnosis * Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=ASDdata, REML = FALSE)
summary(Bmodel3_int)

#MOT_mlu
powerCurveInt_MOTmlu = powerCurve(Bmodel3_int,fixed("MOT_MLU"), along="Child.ID", nsim = 50)
plot(powerCurveInt_MOTmlu)
powerCurveInt_MOTmlu

#ADOS1
powerCurveInt_ADOS = powerCurve(Bmodel3_int,fixed("ADOS1"), along="Child.ID", nsim = 50)
plot(powerCurveInt_ADOS)
powerCurveInt_ADOS

#tokens_MOT
powerCurveInt_tMOT = powerCurve(Bmodel3_int,fixed("tokens_MOT"), along="Child.ID", nsim = 50)
plot(powerCurveInt_tMOT)
powerCurveInt_tMOT

#tokens_CHI
powerCurveInt_tCHI = powerCurve(Bmodel3_int,fixed("tokens_CHI"), along="Child.ID", nsim = 50)
plot(powerCurveInt_tCHI)
powerCurveInt_tCHI

#interaction diagnosis:Visit
powerCurveInt_DiagVisit = powerCurve(Bmodel3_int,fixed("Diagnosis:Visit"), along="Child.ID", nsim = 50)
plot(powerCurveInt_DiagVisit)
powerCurveInt_DiagVisit
```


- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
```{r}

```

- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}


### Riccardo's clumsy function to simulate new participants
### TO DO points are only notes for myself, so not part of the assignment

createNewData <- function (participants,visits,model){
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(Visit=1:visits,Child.ID=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$Child.ID]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma=cov(ranef(model)$Child.ID))
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$Visit + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}
```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.

```{r}

# make subset of 30 kids, 15 with each diagnosis. I could not make Riccardos function work. I decided to count them instead. I do apologise :(
thirty_kids = subset(ASDdata[which(ASDdata$Child.ID>30) ,])


#Assessing power. The smaller the effect size, the more participants we need to get adequate power
Model_30kids = lmer(CHI_MLU ~ Diagnosis + Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=thirty_kids, REML = FALSE)
summary(Model_30kids)

#Power for visit = 100%
#effect size = 0.10
powerKids <- powerSim(Model_30kids,fixed("Visit"), nsim=50)
powerKids


#Power for diagnosis 10%
powerKids_Diag <- powerSim(Model_30kids,fixed("Diagnosis"),nsim=50)
powerKids_Diag

#Power MOT MLU = 100%
#Effect size for MOT_MLU is = 0.30
powerKids_MOTMLU <- powerSim(Model_30kids, fixed("MOT_MLU"), nsim = 50)
powerKids_MOTMLU

#Power for ADOS = 8%
#Effect size for ADOS1 is = -0.013 
powerKids_ADOS1 <- powerSim(Model_30kids, fixed("ADOS1"), nsim = 50)
powerKids_ADOS1

#Power for tokens_mot 62%
#Effect size for tokens_MOT is -0.00023
powerKids_tokens_MOT <- powerSim(Model_30kids, fixed("tokens_MOT"), nsim = 50)
powerKids_tokens_MOT

#Power for tokens_chi 100%
#Effect size for tokens_CHI is 0.0017
powerKids_tokens_CHI <- powerSim(Model_30kids, fixed("tokens_CHI"), nsim = 50)
powerKids_tokens_CHI


#Interaction model_______________________________________________________________________________
#Assessing power. The smaller the effect size, the more participants we need to get adequate power
ModelInt_30kids = Bmodel3 <- lmer(CHI_MLU ~ Diagnosis * Visit + MOT_MLU + ADOS1 + tokens_MOT + tokens_CHI + (1+Visit|Child.ID), data=thirty_kids, REML = FALSE)
summary(ModelInt_30kids)

#Power for interaction diagnosis:visit = 100%
#effect size = 0.10
powerKidsInt <- powerSim(ModelInt_30kids,fixed("Diagnosis:Visit"), nsim=50)
powerKidsInt

#Power MOT MLU = 100%
#Effect size for MOT_MLU is = 0.38
powerKidsInt_MOTMLU <- powerSim(ModelInt_30kids, fixed("MOT_MLU"), nsim = 50)
powerKidsInt_MOTMLU

#Power for ADOS = 8%
#Effect size for ADOS1 is = -0.013 
powerKidsInt_ADOS1 <- powerSim(ModelInt_30kids, fixed("ADOS1"), nsim = 50)
powerKidsInt_ADOS1

#Power for tokens_mot 62%
#Effect size for tokens_MOT is -0.00023
powerKidsInt_tokens_MOT <- powerSim(ModelInt_30kids, fixed("tokens_MOT"), nsim = 50)
powerKidsInt_tokens_MOT

#Power for tokens_chi 100%
#Effect size for tokens_CHI is 0.0017
powerKidsInt_tokens_CHI <- powerSim(ModelInt_30kids, fixed("tokens_CHI"), nsim = 50)
powerKidsInt_tokens_CHI





```




